Everything you wanted to know about PanDA but were truly truly afraid to ask

### 2017-03-09 overall.tex
Iteration section V

% -----------------------------------------------------------------------------
% \subsection{Geant4 Tasks Performance Analysis on Titan}
% \label{ssec:panda_titan}

Currently, 20 instances of PanDA Broker are deployed on a total of 4 DTNs, 5
instances for each DTN. Each broker controls from 15 to 300 detector simulation
jobs per submission, for a theoretical maximum concurrent use of 96,000
cores. Since November 2015, the PanDA Brokers have operated only in backfill
mode, without a defined time allocation, and running at the lowest priority on
Titan.

Figure~\ref{fig:core-hours-utilization} shows Titan core hours used by ATLAS
from January 2016 to February 2017. During that period, ATLAS consumed a total
of 73.8 million Titan core hours, for an average of 7M hours a month, with a
minimum of 3.3M hours in April 2016 and a maximum 14.8M hours in February 2017.
On February, PanDA Brokers used almost twice as much backfill availability than
in any other month. This is likely due to the hardware upgrades made to the
DTNs. The absence of continuous monitoring of the DTNs does not allow for the
specific identification of the bottleneck but spot measurements of the DTNs load
indicate that a faster CPU and better networking were likely responsible for the
improved performance. No relevant code update was made between January and
February 2017 and PanDA Brokers logs indicated that they were able to respond
more promptly to backfill availability.

\begin{figure}[htp]
    \subfloat[]{
        \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
        \label{fig:core-hours-utilization}
    }

    \subfloat[]{
        \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
        \label{fig:backfill-utilization}
    }
\caption{(a).(b).}
\end{figure}

Investigations of the CPU load of the upgraded DTNs shows an average
XX\mtnote{ask Danila for data} load. This indicates that further hardware
upgrades would not be likely to improve significantly the performance of the
PanDA Brokers. Nonetheless, the current load suggests that the number of brokers
per DTN could be increased. This would enable the submission of a larger number
of concurrent jobs to the Titan's PBS queue, allowing for PanDA to consume a
higher percentage of the overall backfill availability.

Figure~\ref{fig:backfill-utilization} shows backfill utilization efficiency,
defined as the fraction of Titan’s total cores available via backfill utilized
by ATLAS, from January 2016 to February 2017. ATLAS reached 18\% average
efficiency with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\%
efficiency on February 2017. The number of total backfill cores available in
April 2016 was 38.1M and in February 2017 33.1M. This shows that the improvement
in the efficiency was determine by the different in total backfill availability.

Figure~\ref{fig:hpc-workload-utilization} shows that between January 2016 and
February 2017, about XX million detector simulation jobs were completed on
Titan, for a total of XXXM events processed. This is equivalent to XX\% of the
total amount of detector simulations performed by ATLAS in the same period of
time, and XX\% of the total number of events processed. Comparatively, Titan
contributed between XX\% and XX\% of the total WLCG availability. When
accounting for the amount of unused backfill availability and the rate of
improvement of PanDA efficiency, these figures confirms the fundamental role
that supercomputers' resources can play for the LHC Run 3.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
\caption{\mtnote{Placeholder for the diagram/data I asked to Danila and
Sergey}}
\label{fig:hpc-workload-utilization}
\end{figure}

The scalability of PanDA Broker depends on the DTNs' resources. Each PanDA
Broker can execute 1 job on up to 300 work nodes for a total of 4,800 cores.
Scaling above this threshold requires instantiating multiple PanDA Brokers on
the same DTN and, once the DTN resources are saturated, on multiple DTNs. In
principle, this is not a problem on Titan as DTNs are supposed to be available
and can be periodically upgraded. Assuming the efficiency reached on February as
reference, using all the available backfill resources on Titan will require to
instantiate around 60 PanDA Brokers. It is not clear whether this is within the
scope of the current DTN capabilities, especially when considering the increased
staging in and out of files from the DTNs and ATLAS Grid sites.

The current design and architecture of the PanDA Broker proven to be very
robust. The failure rate of jobs produced by issues with PanDA Broker and PanDA
in general is XX\%. This confirms the benefits given by reusing most of the code
base of the PanDA Pilot and adopting RADICAL-SAGA for communicating with Titan's
PBS batch system and monitoring jobs execution. During XX months of executing
production jobs, most of the failures were due to: \ldots, \ldots, and \ldots.
The usual corrective measure were put in place and the progressive decline of
the failure rate confirms that PanDA Broker reached production-grade stability
on Titan.\mtnote{I requested figures about failure types and rate to Sergey and
Danila.}

@@

>>>>

Currently, 20 instances of PanDA Broker are deployed on a total of 4 DTNs, 5
instances for each DTN. Each broker controls from 15 to 300 detector simulation
jobs per submission, one simulation job for each Titan's work node and a
theoretical maximum concurrent use of 96,000 cores. Since November 2015, the
PanDA Brokers have operated only in backfill mode, without a defined time
allocation, and running at the lowest priority on Titan.

We evaluated the efficiency, scalability and reliability of the PanDA Brokers
deployed on Titan along two dimensions: (i) amount of backfill availability
utilized by the PanDA Broker; and (ii) amount of jobs walltime spent performing
detector simulations. We based these evaluations on measuring the number of
cores utilized by ATLAS on Titan, the overall backfill availability on Titan,
the number of detector simulations performed on Titan and on all the resources
available to ATLAS, and the failure rate of these simulations. All the
measurements were performed between January 2016 and February 2017.

% -----------------------------------------------------------------------------
\subsection{Efficiency and Scalability of PanDA Broker on Titan}
\label{ssec:panda_titan}

Figure~\ref{fig:core-hours-utilization} shows the number of core hours used by
ATLAS on Titan, from January 2016 to February 2017. During that period, ATLAS
consumed a total of 73.8M core hours, for an average of 7M hours a month, with a
minimum of 3.3M hours in April 2016 and a maximum 14.8M hours in February 2017.

% \begin{figure}[htp]
%     \subfloat[]{
%         \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
%         \label{fig:core-hours-utilization}
%     }
%
%     \subfloat[]{
%         \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
%         \label{fig:backfill-utilization}
%     }
% \caption{(a).(b).}
% \end{figure}

\begin{figure}[htp]
\includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
\caption{Caption.}
\label{fig:core-hours-utilization}
\end{figure}

On February 2017, PanDA Brokers used almost twice as much backfill availability
than in any other month. No relevant code update was made during that period and
logs indicated that the brokers were able to respond more promptly to backfill
availability. This is likely due to hardware upgrades on the DTNs. The absence
of continuous monitoring of the DTNs does not allow to quantify bottlenecks but
spot measurements of the DTNs load indicate that a faster CPU and better
networking were likely responsible for the improved performance.

Investigations of the CPU load of the upgraded DTNs shows an average
XX\mtnote{asked Danila for data} load. As such, further hardware upgrades seem
unlikely to improve significantly the performance of PanDA Brokers. Nonetheless,
the current load suggests that the number of brokers per DTN could be increased.
This would enable the submission of a larger number of concurrent jobs to
Titan's PBS queue, allowing for PanDA to consume a higher percentage of
backfill availability.

Figure~\ref{fig:backfill-utilization} shows backfill utilization efficiency from
January 2016 to February 2017. Efficiency is defined as the fraction of Titan’s
cores available via backfill that was utilized by the PanDA Brokers. The brokers
reached 18\% average efficiency with a minimum 8.9\% efficiency on April 2016
and a maximum 33.5\% efficiency on February 2017. The number of total backfill
cores available was 38.1M in April 2016, and 33.1M in February 2017. This shows
that the difference in total backfill availability was not responsible for the measured difference in efficiency.

\begin{figure}[htp]
\includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
\caption{Caption.}
\label{fig:backfill-utilization}
\end{figure}

Figure~\ref{fig:hpc-workload-utilization} shows that between January 2016 and
February 2017, about 2.25M million detector simulation jobs were completed on
Titan, for a total of 225M events processed. This is equivalent to 9\% of all
the 250M detector simulations performed by ATLAS in the same period of time, and
3.5\% of the 6.6B events processed by those jobs. Comparatively, Titan
contributed 3.9\% of the total of around 200K cores available to ATLAS on Grid,
Cloud, and HPC infrastructures together. When accounting for the amount of
unused backfill availability and the rate of improvement of PanDA efficiency on
Titan, these figures confirms the relevance of supercomputers' resource
contribution to the LHC Run 3.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
\caption{\mtnote{Placeholder for the diagram/data I asked to Danila and
Sergey. Consider a table instead of a figure.}}
\label{fig:hpc-workload-utilization}
\end{figure}

Detector simulations executed on Titan, all process 100 events. This number has
been derived empirically by observing the average walltime of backfill
resources: too many events would \ldots\mtnote{I need to speak further to Sergey
to clarify this.}. Usually, detector simulations performed on the Grid process a
higher number of events than those performed on Titan. This explains the
difference between the percentage of detector simulations performed on Titan
(9\%) and of events computed on Titan for those simulations (3.9\%). PanDA Event
Service~\cite{panda-event-service} would address this limitation: each PanDA
Broker would be able to pull a dynamic number of events for each simulation,
depending on the available walltime. This functionality is currently being
implemented.

Given the current design and implementation, the utilization of a larger portion
of Titan's backfill availability depends on the the number of concurrent PanDA
Brokers instantiated on the DTNs.
% The number of concurrent detector simulation jobs that PanDA Broker can manage
% depends on the DTNs' resources. Each PanDA Broker can execute 1 MPI PBS job on
% up to 300 work nodes for a total of 4,800 cores. Scaling above this threshold
% requires instantiating multiple PanDA Brokers on the same DTN and, once the
% DTN resources are saturated, on multiple DTNs.
In principle, this is not a problem on Titan as multiple DTNs are supposed to be
available and periodically upgraded. Assuming the efficiency reached on February
as reference, using all the available backfill resources on Titan would require
around 60 concurrent PanDA Brokers. Further testing and better monitoring
capabilities are required to establish whether the current DTN capabilities can
support this load, especially when considering the increased volume of files
staged in and out the DTNs.

% -----------------------------------------------------------------------------
\subsection{Efficiency and Scalability of Detector Simulation on Titan}
\label{ssec:panda_titan}

@@

% ----------------------------------------------------------------------------
\subsection{Reliability of PanDA Broker and Detector Simulations on Titan}
\label{ssec:reliability}

The current design and architecture of the PanDA Broker is proving to be as
reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
Brokers were responsible for around the 19\% of the failures, compared to the
29\% of failures produced by the JobDispatcher module of the PanDA Server, and
the 13\% failures produced by Geant4. The current failure rate confirms the
benefits given by reusing most of the code base of the PanDA Pilot for
implementing the PanDA Broker. It also shows that adopting third-party libraries
like RADICAL-SAGA did not have a measurable adverse effect on reliability.

Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
recorded via the PanDA Brokers between January 2016 and February 2017. The
initial preponderance of unknown failures was progressively addressed by a more
accurate failure mode developed for the PanDA Broker and Titan. It also shows
the progressive improvement of the broker's reliability: In Februrary 2017, most
of the failures were produced by the initialization and configuration phases of
AthenaMP or by Geant4 failed executions.\mtnote{If we want to keep this diagram,
we will need to plot it from the CSV to make it readable.}

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/failures_titan.png}
\caption{Caption.}
\label{fig:failures-titan}
\end{figure}
